---
title: "Prediction-Assignment-Writeup"
author: "TSAI  YI- FAN"
date: "2018/4/22/"
output: html_document
---
Background
=======
Using devices such as JawboneUp, NikeFuelBand, and Fitbitit is now possible to collect a large amount of data about personal activity relatively inexpensively. These type of devices are part of the quantified self movement - a group of enthusiasts who take measurements about themselves regularly to improve their health, to find patterns in their behavior, or because they are tech geeks. One thing that people regularly do is quantify how much of a particular activity they do, but they rarely quantify how well they do it.

In this project, the goal is to use data from accelerometers on the belt, forearm, arm, and dumbell of 6 participants. They were asked to perform barbell lifts correctly and incorrectly in 5 different ways. More information is available from the website: http://groupware.les.inf.puc-rio.br/har (see the section on the Weight Lifting Exercise Dataset).

Preparing the data and R packages
============
oad packages, set caching

```{r echo = TRUE}
require(caret)
require(corrplot)
require(Rtsne)
require(xgboost)
require(stats)
require(knitr)
require(ggplot2)
knitr::opts_chunk$set(cache=TRUE)
 
```

For fast and accurate training the model, I choose XGBoost, an implementation of tree-based extreme gradient boosting algorithm. (As discussed in the course's forum, this XGBoost tool is confirmed by course's CTA to be allowed to be used in this assignment project.)

Getting Data

```{r echo = TRUE}
# URL of the training and testing data
train.url ="https://d396qusza40orc.cloudfront.net/predmachlearn/pml-training.csv"
test.url = "https://d396qusza40orc.cloudfront.net/predmachlearn/pml-testing.csv"
train.name = "./data/pml-training.csv"
test.name = "./data/pml-testing.csv"
if (!file.exists("./data")) {
  dir.create("./data")
} 
if (!file.exists(train.name)) {
  download.file(train.url, destfile=train.name, method="curl")
}
if (!file.exists(test.name)) {
  download.file(test.url, destfile=test.name, method="curl")
}
train = read.csv("./data/pml-training.csv")
test = read.csv("./data/pml-testing.csv")
dim(train)
dim(test)
names(train)
```

The raw training data has 19622 rows of observations and 158 features (predictors). Column X is unusable row number. While the testing data has 20 rows and the same 158 features. There is one column of target outcome named classe.

Data cleaning
First, extract target outcome (the activity quality) from training data, so now the training data contains only the predictors (the activity monitors).

```{r echo = TRUE}
outcome.org = train[, "classe"]
outcome = outcome.org 
levels(outcome)

```

Outcome has 5 levels in character format.
Convert the outcome to numeric, because XGBoost gradient booster only recognizes numeric data.

```{r echo = TRUE}
num.class = length(levels(outcome))
levels(outcome) = 1:num.class
head(outcome)
```

The outcome is removed from training data.

```{r echo = TRUE}
train$classe = NULL
 
```

The assignment rubric asks to use data from accelerometers on the belt, forearm, arm, and dumbell, so the features are extracted based on these keywords.

```{r echo = TRUE}
filter = grepl("belt|arm|dumbell", names(train))
train = train[, filter]
test = test[, filter]
 
```

Instead of less-accurate imputation of missing data, remove all columns with NA values.

```{r echo = TRUE}
cols.without.na = colSums(is.na(test)) == 0
train = train[, cols.without.na]
test = test[, cols.without.na]
 
```

Preprocessing
Check for features's variance
Based on the principal component analysis PCA, it is important that features have maximum variance for maximum uniqueness, so that each feature is as distant as possible (as orthogonal as possible) from the other features.

```{r echo = TRUE}
zero.var = nearZeroVar(train, saveMetrics=TRUE)
zero.var
 
```
There is no features without variability (all has enough variance). So there is no feature to be removed further.

Plot of relationship between features and outcome
Plot the relationship between features and outcome. From the plot below, each features has relatively the same distribution among the 5 outcome levels (A, B, C, D, E).

```{r echo = TRUE}
featurePlot(train, outcome.org, "strip")
 
```

Plot of correlation matrix
Plot a correlation matrix between features.
A good set of features is when they are highly uncorrelated (orthogonal) each others. The plot below shows average of correlation is not too high, so I choose to not perform further PCA preprocessing.

```{r echo = TRUE}
corrplot.mixed(cor(train), lower="circle", upper="color", 
               tl.pos="lt", diag="n", order="hclust", hclust.method="complete")
 
```

tSNE plot
A tSNE (t-Distributed Stochastic Neighbor Embedding) visualization is 2D plot of multidimensional features, that is multidimensional reduction into 2D plane. In the tSNE plot below there is no clear separation of clustering of the 5 levels of outcome (A, B, C, D, E). So it hardly gets conclusion for manually building any regression equation from the irregularity.

```{r echo = TRUE}
tsne = Rtsne(as.matrix(train), check_duplicates=FALSE, pca=TRUE, 
              perplexity=30, theta=0.5, dims=2)
embedding = as.data.frame(tsne$Y)
embedding$Class = outcome.org
g = ggplot(embedding, aes(x=V1, y=V2, color=Class)) +
  geom_point(size=1.25) +
  guides(colour=guide_legend(override.aes=list(size=6))) +
  xlab("") + ylab("") +
  ggtitle("t-SNE 2D Embedding of 'Classe' Outcome") +
  theme_light(base_size=20) +
  theme(axis.text.x=element_blank(),
        axis.text.y=element_blank())
print(g)
 
```

Build machine learning model
Now build a machine learning model to predict activity quality (classe outcome) from the activity monitors (the features or predictors) by using XGBoost extreme gradient boosting algorithm.

XGBoost data
GBoost supports only numeric matrix data. Converting all training, testing and outcome data to matrix.

```{r echo = TRUE}
# convert data to matrix
train.matrix = as.matrix(train)
mode(train.matrix) = "numeric"
test.matrix = as.matrix(test)
mode(test.matrix) = "numeric"
# convert outcome from factor to numeric matrix 
#   xgboost takes multi-labels in [0, numOfClass)
y = as.matrix(as.integer(outcome)-1)

```

XGBoost parameters
Set XGBoost parameters for cross validation and training.
Set a multiclass classification objective as the gradient boosting's learning function.
Set evaluation metric to merror, multiclass error rate.

```{r echo = TRUE}
# xgboost parameters
param <- list("objective" = "multi:softprob",    # multiclass classification 
              "num_class" = num.class,    # number of classes 
              "eval_metric" = "merror",    # evaluation metric 
              "nthread" = 8,   # number of threads to be used 
              "max_depth" = 16,    # maximum depth of tree 
              "eta" = 0.3,    # step size shrinkage 
              "gamma" = 0,    # minimum loss reduction 
              "subsample" = 1,    # part of data instances to grow tree 
              "colsample_bytree" = 1,  # subsample ratio of columns when constructing each tree 
              "min_child_weight" = 12  # minimum sum of instance weight needed in a child 
              )
 
```

Expected error rate
Expected error rate is less than 1% for a good classification. Do cross validation to estimate the error rate using 4-fold cross validation, with 200 epochs to reach the expected error rate of less than 1%.

```{r echo = TRUE}
set.seed(1234)
nround.cv = 200
system.time( bst.cv <- xgb.cv(param=param, data=train.matrix, label=y, 
              nfold=4, nrounds=nround.cv, prediction=TRUE, verbose=FALSE) )
```

Elapsed time is around 150 seconds (2.5 minutes).

```{r echo = TRUE}
tail(bst.cv$dt) 
 
```

From the cross validation, choose index with minimum multiclass error rate.
Index will be used in the model training to fulfill expected minimum error rate of < 1%.

```{r echo = TRUE}
# get CV's prediction decoding
pred.cv = matrix(bst.cv$pred, nrow=length(bst.cv$pred)/num.class, ncol=num.class)
pred.cv = max.col(pred.cv, "last")
# confusion matrix
confusionMatrix(factor(y+1), factor(pred.cv))
```

```{r echo = TRUE}

 
```

```{r echo = TRUE}


```